---
title: "INFO523-R-HW1-KashyapNadendla"
format:
  html:
    theme: default
    author: "Kashyap Sai Prasad Nadendla"
---

## SETUP

```{r setup}

#install.packages("pacman")

library(pacman)

p_load(dlookr,
       DMwR2, # Data Mining with R functions
       GGally, # Pair-wise plots using ggplot2
       Hmisc, # Data analysis 
       palmerpenguins, # Alternative to the Iris dataset
       tidyverse) # Data wrangling, manipulation, visualization
```

```{r}

data(algae, package = "DMwR2")

algae |> glimpse()
```

# **Central tendency: mean, median, mode**

## Mean

```{r}

algae$a1 |>
  mean()
```

## Median

```{r}

algae$a1 |>
  median()
```

## Mode

```{r}

Mode <- function(x,na.rm = FALSE){
  if(na.rm) x<-x[!is.na(x)]
  ux <- unique(x)
  return (ux[which.max(tabulate(match(x, ux)))])
}

algae$a2 |> Mode()
```

## DMwR centralValue() function:

```{r}

algae$a1 |> centralValue()
```

```{r}

algae$speed |> centralValue()
```

# **Statistics of spread (variation)**

## **Variance**

```{r}

algae$a1 |> var()
```

## **Standard deviation**

```{r}

algae$a1 |> sd()
```

## **Range**

```{r}

algae$a1 |> range()
```

## **Maximum value**

```{r}

algae$a1 |> max()
```

## **Minimum value**

```{r}

algae$a1 |> min()
```

## **Interquartile range**

```{r}

algae$a1 |> IQR()
```

## **Quantiles**

```{r}

algae$a1 |> quantile()
```

Specifying specific quantiles:

```{r}

algae$a1 |> quantile(probs = c(0.2, 0.8))
```

# **Missing values**

```{r}

library(purrr)

nas <- algae %>% 
  purrr::map_dbl(~sum(is.na(.))) %>% 
  sum()

cat("The dataset contains ", nas, "NA values. \n")
```

```{r}

incomplete_rows <- algae %>% 
  summarise_all(~!complete.cases(.)) %>%
  nrow()

cat("The dataset contains ", incomplete_rows, "(out of ", nrow(algae),") incomplete rows. \n")
```

# **Summaries of a dataset**

## Base R's summary()

```{r}

algae |> summary()
```

## Himsc's describe()

```{r}

data("penguins")
penguins |> Hmisc::describe()

```

## dlookr's describe()

```{r}

penguins |> dlookr::describe()
```

# **Summaries on a subset of data**

```{r}

algae |>
  summarise(avgNO3 = mean(NO3, na.rm=TRUE),
            medA1 = median(a1))
```

```{r}

algae |>
  select(mxPH:Cl) |>
  summarise_all(list(mean, median), na.rm = TRUE)

```

```{r}

algae |>
  select(a1:a7) |>
  summarise_all(funs(var))
```

```{r}

algae |>
  select(a1:a7) |>
  summarise_all(c("min", "max"))
```

## Use summarise() with group_by()

```{r}

algae |>
  group_by(season, size) |>
  summarise(nObs = n(), mA7 = median(a7))
```

```{r}

penguins |> 
  group_by(species) |>
  summarise(var = var(bill_length_mm, na.rm = TRUE))
```

## **Aggregating data**

```{r}

penguins |>
  group_by(species) |>
  reframe(var = quantile(bill_length_mm, na.rm = TRUE))
```

```{r}

```

```{r}

# Note: Aggregating data with summarize was deprecated in dplyr 1.1.0, reframe() should be used instead. Use dlookr…

penguins |>
  group_by(species) |>
  dlookr::describe(bill_length_mm)

```
